Unified Convergence Analysis of FedAve on Non-IID Data

### 2020-05-23 source/save.tex
first draft

% !TEX ROOT=./main.tex

\section{Convergence Results}
This section summarizes the current results and assumptions.
\begin{table}[h!]
\centering
\small
	\begin{tabular}{|c|c|c|c|c|}\hline
		paper         &  Cvx Non-smth & Cvx L-smth & Strongly Cvx Non-smth& Strongly Cvx L-smth \\ \hline
	Accelerated	SGD   &   $\cO(\frac{1}{\sqrt{T}})$     & $\cO(\frac{1}{\sqrt{T}})$   &    $\cO(\frac{1}{T})$   &  $\cO(\frac{1}{KT})$    \\\hline
	SGD    &    $\cO(\frac{1}{\sqrt{T}})$  &   $\cO(\frac{1}{\sqrt{T}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}, $\cO(\frac{1}{KT})$ ours    \\\hline
	\end{tabular}
	\caption{Summarize of related work on federated learning, heterogeneous data.}
\end{table}

Linear speedup of SGD and MaSS for linear regression, full participation: 
\begin{align*}
\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N}+(E-1)^{2})(\tilde{\kappa}_{N}+(E-1)^{2})}})
\end{align*}

 which implies that for $E=1$, there is linear speedup, i.e. 
\begin{align*}
\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N})(\tilde{\kappa}_{N})}})\leq C\cdot\exp(-\frac{Nt}{\sqrt{(\kappa_{1})(\tilde{\kappa})}})
\end{align*}
when $N$ satifies $N=O(\min(L_{1}/L,\tilde{\kappa}))$, and for $(E-1)^{2}\leq\min(\kappa_{N},\tilde{\kappa}_{N})$,
\begin{align*}
\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{Nt}{4\sqrt{\kappa_{1}\tilde{\kappa}}})
\end{align*}

Acceleration of MaSS for linear regression, partial participation $O((1+\frac{4}{K}\cdot(\frac{t}{E}-1))\exp(-\frac{t}{\sqrt{\kappa_{m}\tilde{\kappa}_{m}}}))$
over $O((1+\frac{4}{K}\cdot(\frac{t}{E}-1))\exp(-\frac{t}{\kappa_{m}}))$.

Linear speedup of SGD and Nesterov SGD for strongly convex and smooth
objectives: full participation $O(\frac{\nu_{max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{2}})$
with $\nu_{max}=N\cdot\max_{k}p_{k}$; partial participation $O(\frac{\nu_{max}^{2}\sigma^{2}+4E^{2}G^{2}L}{KT})$.
In particular, this implies full participation allows $E=O(\sqrt{\frac{T}{N}})$
to achieve $O(1/NT)$, whereas partial participation only allows $E=O(1)$
to achieve $O(1/KT)$.

Other accelerations likely not possible.

\begin{assumption}[Bounded local variance]
Let $\xi_{t}^{k}$ be sampled from the $k$ -th device's local data uniformly at random. The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
for $k=1, \cdots, N$.	
\end{assumption}

% \begin{assumption}
% The expected squared norm of stochastic subgradients is uniformly bounded. i.e.,
% $\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.  This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$. Denote $G^2 = \sum_{k=1}^N p_k G_k^2$
% \label{ass:subgrad2}
% \end{assumption}

\begin{assumption}[Expected square norm of gradient]
The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$. This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
\label{ass:subgrad2}
\end{assumption}

\begin{assumption}[L-smooth]
$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
\label{ass:lsmooth}
\end{assumption}
\begin{assumption}[Strongly-convex]
$	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}-\mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
\label{ass:stroncvx}
\end{assumption}

\begin{definition}[Squared Weighted Gradient Diversity]
	We indicate the following quantity as weighted gradient diversity among the local 
	objectives:
	\begin{align*}
		\frac{\sum_{k=1}^K p_k  \mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}, \xi^{k}\right)\right\|^{2} }{\mathbb{E}\left\| \sum_{k=1}^K p_k   \nabla F_{k}\left(\mathbf{w}, \xi^{k}\right)\right\|^{2} } \leq \lambda 
	\end{align*}
\end{definition}
This definition is similar to the weighted gradient diversity defined in~\cite{haddadpour2019convergence}.

\begin{definition}[Gradient Dissimilarity~\cite{li2018federated}]
	We indicate the following quantity as gradient diversity among the local 
	objectives:
	\begin{align*}
		\frac{\sum_{k=1}^K p_k \left\|\nabla F_{k}\left(\mathbf{w}\right)\right\|^{2} }{\left\|   \nabla F\left(\mathbf{w}\right)\right\|^{2} } \leq \lambda 
	\end{align*}
	For some $\epsilon>0$, for all the points $\vw \in \mathcal{S}_{\epsilon}^{c}=\left\{\vw |\|\nabla F(\vw)\|^{2}>\epsilon\right\}$
\end{definition}

\section{The Convergence of FedAve for General Convex Functions}
This section summarizes the convergence of FedAve.
The updates of FedAve with partial device activation is given by: 
\begin{align} 
\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
\end{align}

>>>>

\section{The Convergence of FedAve for General Convex Functions}

### 2020-05-23 source/sec_exp.tex
first draft

\begin{figure}[h!]
\centering
\begin{tabular}{cc}
	\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupNodesT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} & 
\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} 
\end{tabular}
	\caption{Logistic regression with regularization $1e-5$. The linear speed up w.r.t the number of nodes and number of epochs. The synthetic dataset has $N=49749$ samples, evenly distributed on $1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1024$ devices. The figure shows the number of iterations/rounds needed to converge to $\epsilon-$accuracy. The learning rate is decayed as the $\eta_t = \min(32, \frac{Nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c, 2^{-2}c, c, 2c, 2^{2}c\}$ and for each configuration. Batch size is 4 in this case.}
\end{figure}

>>>>

\begin{figure}[h!]
\centering
\begin{tabular}{cc}
	\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupNodesT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} & 
\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} \\
	\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupNodesRounds-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} & 
\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsRounds-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} \\
\end{tabular}
	\caption{Logistic regression with regularization $1e-5$. The linear speed up w.r.t the number of nodes and number of epochs. The synthetic dataset has $N=49749$ samples, evenly distributed on $1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1024$ devices. The figure shows the number of iterations/rounds needed to converge to $\epsilon-$accuracy. The learning rate is decayed as the $\eta_t = \min(32, \frac{Nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c, 2^{-2}c, c, 2c, 2^{2}c\}$ and for each configuration. Batch size is 4 in this case.}
\end{figure}

### 2020-05-23 source/sec_problem.tex
reorg
first draft

\section{Problem Setting}

\begin{align}
	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\}
	\label{eq:problem}
\end{align}
where $N$ is the number of devices, and $p_k$ is the weight of the k-th device
such that $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. Suppose the k-th device
holds the nk training data: $x_{k,1}, x_{k,2}, \dots, x_{k,n_k}$ . The local
$k \in S_t^K$ objective $F_k(\cdot)$ is defined by

\begin{align}
F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; x_{k, j}\right)	
\label{eq:localloss}
\end{align}

% The summary of all assumptions:

% \begin{assumption}
% $F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
% \end{assumption}

% \begin{assumption}
% $F_1,\dots, F_N$ are all convex: for all $\vv$ and $\vw$, 
% $F_k(\vv) \geq F_k(\vw)+(\vv -\vw)^T \grad F_k(\vw)$. \label{ass:cvx}
% \end{assumption}
% \begin{assumption}
% $	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
% \end{assumption}

% \begin{assumption}
% Let $\xi_{t}^{k}$ be sampled from the k-th device's local data uniformly at random. The variance of stochastic gradients in each device is bounded:
% $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$, for $k = 1,..., N$.
% \end{assumption}

% \begin{assumption}
% The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
% $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
% \end{assumption}

\section{Notations}

Let $\vw_t^k$ be the model parameter maintained in the k-th device at the t-th step. $\cI_E$ is a set of global synchronization steps, i.e., $\cI_E = \{n E|n = 1, 2, 3,\dots \}$. If $t+1 \in \cI_E$, which means we communicate the server with (all) clients at time step $t+1$. $\vv^k_{t+1}$ is the immediate result of one step SGD update from $\vw^k_{t}$.
$\xi_{t}^{k}$ denotes the data sampled from k-th device’s local data uniformly at random.
Follow the common practice, we define two virtual sequences $\overline{\mathbf{v}}_{t}$ and $\overline{\mathbf{w}}_{t}$. For full device participation and $t \notin \cI_E$,
$\ov{v}_t = \ov{w}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$. In partial participation, $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\sum_{k\in \cS_t}\mathbf{w}_{t}^{k}$. However, we can
set unbiased sampling strategy such that $ \EE_{\cS_t} \ov{w}_t = \ov{v}_t$.
$\overline{\mathbf{v}}_{t+1}$ is one-step SGD from $\overline{\mathbf{w}}_{t}$. 
\begin{align}
\overline{\mathbf{v}}_{t+1}=\overline{\mathbf{w}}_{t}-\eta_{t} \mathbf{g}_{t}	\label{eq:vbar}
\end{align}
% $t+1 \in \cI_E$, we can fetch $\overline{\vw}_{t+1}$, we can communicate $\overline{\vw}_{t+1}$ to all devices.
where $\vg_{t} = \sum_{k=1}^{N} p_{k} \vg_{t,k} $ is one-step stochastic gradient, averaged over all devices. 
\begin{align}
\vg_{t,k} \left\{\begin{array}{ll} 
 = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right)  &  \text{smooth}\\
 \in \partial F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right)  & \text{non-smooth}
 \end{array}\right.
\end{align}
Similarly, we denote the expected one-step gradient $\ov{g}_{t}= \EE_{\xi_t}[\vg_t] = \sum_{k=1}^{N} p_{k} \EE_{\xi_{t}^{k}} \vg_{t,k}$, where
\begin{align}
\EE_{\xi_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
 = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)  &  \text{smooth}\\
 \in \partial F_{k}\left(w_{t}^{k}\right)  & \text{non-smooth}
 \end{array}\right.
\end{align}
and we use $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes samples at all devices at time step $t$.

The updates of FedAve with partial device activation is given by: 
\begin{align} 
\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
\sum_{k \in \cS_{t+1}} p_{k} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
\end{align}

In~\cite{li2019convergence}, two types of unbiased sampling strategies are considered in Lemma 5. 
The sampling scheme I establishes $\cS_{t+1}$ by i.i.d. sampling the devices with replacement,
in this case the upper bound of expected square norm of $\ov{w}_{t+1} - \ov{v}_{t+1}$ is given by:
\begin{align}
\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4}{K} \eta_t^2 E^2G^2
\end{align}
The sampling scheme II establishes $\cS_{t+1}$ by uniformly sampling all devices without
replacement, in which we have the 
\begin{align}
\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4(N - K)}{K(N-1)} \eta_t^2 E^2G^2
\end{align}
We denote this upper bound as follows for concise presentation. 
\begin{align}
	\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2 \leq  \eta_t^2 C
	\label{eq:partialsample}
\end{align}

\lkxcom{quantify biased sampling scheme ?}

>>>>

\section{Problem Setup}

\subsection{The Federated Learning Problem}

The federated learning problem, given local objectives $F_k$ and device weights $p_k$, is to minimize

\begin{align}
	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\}
	\label{eq:problem}
\end{align}

where $N$ is the number of devices, and device weights $p_k$ satisfy $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. Suppose the $k$-th device
holds the $n_k$ data points: $\xi_{k,1}, \xi_{k,2}, \dots, \xi_{k,n_k}$ . We assume that the local objective $F_k(\cdot)$ is defined by

\begin{align}
F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \xi_{k, j}\right)	
\label{eq:localloss}
\end{align}

where $\ell$ is some loss function. Thus, in the federated learning setting, each device stores its local data, which gives rise to its own local objective $F_k$. A key difference with many problems in the distributed optimization setting is that data on local devices is heterogeneous, i.e. $F_k(w)$ may have different minimizers $w^{\ast}_k$. Moreover, each local device only has access to stochastic gradients based on data stored on the local device. In this paper, we investigate local objectives that are convex or strongly convex and smooth, although convergence results for non-smooth objectives are also given in the appendix.

\subsection{The FedAvg Algorithm}

Despite the heterogeneity in local objectives, the federated learning
problem can still be solved efficiently in a distributed manner with
the simple Federated Average (FedAvg) algorithm, which consists of
the iterations

\begin{align}
\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}g_{t,k}\\
\vw_{t+1}^{k} & =\begin{cases}
\vv_{t+1}^{k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\vv_{t+1}^{k} & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}
\end{align}

where $\vw_t^k$ is the local model parameter maintained in the $k$-th device during the $t$-th iteration, 
\begin{align}
g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
\end{align} 
is the stochastic gradient based on $\xi_{t}^{k}$, the data sampled from k-th device’s local data uniformly at random. $\mathcal{I}_{E}=\{E,2E,\dots\}$
is the set of global synchronization steps. Thus when $t+1 \notin \mathcal{I}_{E}$, each device updates its local parameter based on local
stochastic gradient descent, and every $E$ iterations, parameters are aggregated
across devices based on weights $p_{k}$ and then broadcast back to
each local device.

Communication and broadcasting of local parameters
is costly, so the efficiency of the FedAvg algorithm requires that
$E$ is relatively large, while at the same time ensuring convergence
to a global minimizer of the global objective $\sum_{k}p_{k}F_{k}(w)$.

In practical applications, it may not be possible to collect local
parameters from \emph{all }devices at a specified communication time.
Thus the above \emph{full participation }FedAvg algorithm may not
be feasible. Instead, a more practical algorithm is to allow only
a random subset $\mathcal{S}_{t+1}$ of all devices to participate in parameter aggregation, while
broadcasting the aggregated parameter to all devices. This gives rise
to the \emph{partial participation }FedAvg algorithm.

The updates of FedAvg with partial device activation is given by: 
\begin{align} 
\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
\end{align}

where $\mathcal{S}_{t+1}$ is obtained using some sampling scheme. We consider two types of common unbiased sampling strategies. The first one samples each device with probability $p_k$, with replacement, while the second one samples each device uniformly without replacement. Both of these sampling schemes guarantee that updates in FedAvg with partial participation are unbiased stochastic versions of updates in FedAvg with full participation, during the communication round.

\subsection{Notation and Assumptions}

Following common practice, we define two virtual sequences $\overline{\mathbf{v}}_{t}$ and $\overline{\mathbf{w}}_{t}$. For full device participation and $t \notin \cI_E$,
$\ov{v}_t = \ov{w}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$. For partial participation, $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\ov{w}_t=\sum_{k\in \cS_t}\mathbf{w}_{t}^{k}$. However, we can
set unbiased sampling strategy such that $ \EE_{\cS_t} \ov{w}_t = \ov{v}_t$.
$\overline{\mathbf{v}}_{t+1}$ is one-step SGD from $\overline{\mathbf{w}}_{t}$. 
\begin{align}
\overline{\mathbf{v}}_{t+1}=\overline{\mathbf{w}}_{t}-\eta_{t} \mathbf{g}_{t}	\label{eq:vbar}
\end{align}
% $t+1 \in \cI_E$, we can fetch $\overline{\vw}_{t+1}$, we can communicate $\overline{\vw}_{t+1}$ to all devices.
where $\vg_{t} = \sum_{k=1}^{N} p_{k} \vg_{t,k} $ is one-step stochastic gradient, averaged over all devices. 
\begin{align}
\vg_{t,k} \left\{\begin{array}{ll} 
 = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right)  &  \text{smooth}\\
 \in \partial F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right)  & \text{non-smooth}
 \end{array}\right.
\end{align}
Similarly, we denote the expected one-step gradient $\ov{g}_{t}= \EE_{\xi_t}[\vg_t] = \sum_{k=1}^{N} p_{k} \EE_{\xi_{t}^{k}} \vg_{t,k}$, where
\begin{align}
\EE_{\xi_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
 = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)  &  \text{smooth}\\
 \in \partial F_{k}\left(w_{t}^{k}\right)  & \text{non-smooth}
 \end{array}\right.
\end{align}
and $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes samples at all devices at time step $t$.

We will require some standard assumptions on the local objectives and their gradients.

\begin{assumption}[Bounded local variance]
	Let $\xi_{t}^{k}$ be sampled from the $k$-th device's local data uniformly at random. The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$.
\end{assumption}

\begin{assumption}[Expected square norm of gradient]
	The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
	$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
	\label{ass:subgrad2}
\end{assumption}

\begin{assumption}[L-smooth]
	$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
	\label{ass:lsmooth}
\end{assumption}

\begin{assumption}[Strongly-convex]
	$	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}-\mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
	\label{ass:stroncvx}
\end{assumption}

Even though Assumption \ref{ass:subgrad2} also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$, which further implies $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq G^2$, the two assumptions are stated separately because it makes explicit the dependence of the optimality gap on different terms.
